\texttt{nuance}: Detection of planetary transits in the presence of correlated noise

### 2023-12-02 latex/paper.tex
aal figures go in workflows

\begin{figure}[H]
    \begin{centering}
        \includegraphics[width=8.5cm]{issue1.pdf}
        \caption{Illustration of the effect of correlated noise on a single transit SNR. A 1-hour transit signal of depth 1\% is generated on top of white noise (variance of $0.0015^2$) as part of a 24-hours observation with an exposure time of 1 minute (top). Then, in the bottom plot, correlated noise is added, generated using a Gaussian Process with a Matèrn-32 kernel of scale 1 hour and sigma of 0.2\%. The SNR on the right of each light curve is computed using \autoref{eq:snr}. Models used to simulate these data are provided in \autoref{signals_simulations}.}
        \label{fig:issue1}
    \end{centering}
\end{figure}

@@

\begin{figure}[H]
    \begin{centering}
        \includegraphics[width=0.85\linewidth]{issue2.pdf}
        \caption{(Top left) Simulated observation spanning 3 days, including a transit signal with a depth of 0.8\%, duration of 0.1 day, and measurement error of 1.5\% (using the model from \cite{protopapas} described in \autoref{app_proto}). (Left) Correlated noise in the form of stellar variability with different timescales and amplitudes added to the unique transit signal. This signal is then reconstructed using (in purple) Tukey's bi-weight filter \citep{wotan} with an optimal window size of three times the transit duration, and (in red) a Gaussian Process with the same kernel used to simulate the data. In each case the variability is reconstructed, subtracted, and the transit SNR computed using \autoref{eq:snr} with the transit depth estimated as the minimum in-transit flux.}
        \label{fig:issue2}
    \end{centering}
\end{figure}
In order to explore the parameter space for which detrending is the most problematic, we employ the method previously described to simulate 10 000 light curves containing a transit signal and correlated noise in the form of stellar variability. For each light curve, we reconstruct the variability signal and detrend it using an optimal bi-weight filter before estimating the remaining transit SNR. For convenience, the simulated light curves share a common transit added on top of white noise, and a variability signal only defined by two parameters: $\tau$, the relative timescale of the variability with respect to the transit duration; and $\delta$, the relative amplitude of the variability against transit depth, corresponding to an SHO kernel with hyperparameters chosen as: 
\begin{equation}\label{eq:relative_params}
    \omega = \frac{\pi}{\tau D}, \hspace{0.5cm} 
    \sigma = \delta \frac{\Delta}{2} \hspace{0.5cm}  \text{and}  \hspace{0.5cm}  
    Q = 10,
\end{equation}
with $\Delta=1\%$ and $D=0.04$ days, the depth and duration of the simulated transit. 
% We fix a relatively high value for the quality factor $Q$ in order to restrict our simulations to strongly periodic variability signals.

@@

\begin{figure}[H]
    \begin{centering}
        \includegraphics[width=0.8\linewidth]{principle_linear_search.pdf}
        \caption{Principle and output of the \textit{linear search}. The simulated dataset (top) corresponds to the one shown and described in \autoref{fig:app_principle_dataset}. First, a set of durations and depths $\set{T_i, D_j}_{i,j}$ is generated. For each pair of indices $(i,j)$, the likelihood $\ln p(\bm{f} \vert T_i ,D_j, \Delta_{i,j})$ is computed using the parameters from \autoref{eq:ls} and the expression of \autoref{eq:linear_search_ll}. This process yields the grid of log-likelihoods $\ln\mathcal{L}$ (bottom plot), as well as the $\set{\Delta_{i,j}, \sigma_{i,j}}_{i, j}$ transit depths and errors inferred linearly using \autoref{eq:ls}.}
        \label{fig:linear_search}
    \end{centering}
\end{figure}

@@

\begin{figure}[H]
    \centering
    \includegraphics[width=0.8\linewidth]{principle_periodic_0.pdf}\\
    {(a) $\ln \mathcal{L}$}

    \centering
    \includegraphics[width=0.8\linewidth]{principle_periodic_1.pdf}\\
    {(b) $\mathcal{P}(P)$}
\end{figure}

@@

\begin{figure}[H]
    \begin{centering}
        \includegraphics[width=0.8\linewidth]{principle_Q.pdf}
        \caption{For each period $P$, the joint likelihood $\mathcal{P}(P)$ is computed using \autoref{eq:result}, and the value of the maximum likelihood transit SNR retained as $\mathcal{Q}(P)$.}
        \label{fig:periodogram}
    \end{centering}
\end{figure}

>>>>

\begin{figure}[H]
    \begin{centering}
        \includegraphics[width=8.5cm]{../workflows/plot_issues/figures/issue1.pdf}
        \caption{Illustration of the effect of correlated noise on a single transit SNR. A 1-hour transit signal of depth 1\% is generated on top of white noise (variance of $0.0015^2$) as part of a 24-hours observation with an exposure time of 1 minute (top). Then, in the bottom plot, correlated noise is added, generated using a Gaussian Process with a Matèrn-32 kernel of scale 1 hour and sigma of 0.2\%. The SNR on the right of each light curve is computed using \autoref{eq:snr}. Models used to simulate these data are provided in \autoref{signals_simulations}.}
        \label{fig:issue1}
    \end{centering}
\end{figure}

@@

\begin{figure}[H]
    \begin{centering}
        \includegraphics[width=1\linewidth]{../workflows/plot_issues/figures/issue2.pdf}
        \caption{(Top left) Simulated observation spanning 3 days, including a transit signal with a depth of 0.8\%, duration of 0.1 day, and measurement error of 1.5\% (using the model from \cite{protopapas} described in \autoref{app_proto}). (Left) Correlated noise in the form of stellar variability with different timescales and amplitudes added to the unique transit signal. This signal is then reconstructed using (in purple) Tukey's bi-weight filter \citep{wotan} with an optimal window size of three times the transit duration, and (in red) a Gaussian Process with the same kernel used to simulate the data. In each case the variability is reconstructed, subtracted, and the transit SNR computed using \autoref{eq:snr} with the transit depth estimated as the minimum in-transit flux.}
        \label{fig:issue2}
    \end{centering}
\end{figure}
In order to explore the parameter space for which detrending is the most problematic, we employ the method previously described to simulate 10 000 light curves containing a transit signal and correlated noise in the form of stellar variability. For each light curve, we reconstruct the variability signal and detrend it using an optimal bi-weight filter before estimating the remaining transit SNR. For convenience, the simulated light curves share a common transit added on top of white noise, and a variability signal only defined by two parameters: $\tau$, the relative timescale of the variability with respect to the transit duration; and $\delta$, the relative amplitude of the variability against transit depth, corresponding to an SHO kernel with hyperparameters chosen as: 
\begin{equation}\label{eq:relative_params}
    \omega = \frac{\pi}{\tau D}, \hspace{0.5cm} 
    \sigma = \delta \frac{\Delta}{2} \hspace{0.5cm}  \text{and}  \hspace{0.5cm}  
    Q = 10,
\end{equation}
with $\Delta=1\%$ and $D=0.04$ days, the depth and duration of the simulated transit. 
% We fix a relatively high value for the quality factor $Q$ in order to restrict our simulations to strongly periodic variability signals.

@@

\begin{figure}[H]
    \begin{centering}
        \includegraphics[width=0.85\linewidth]{../workflows/principle/figures/principle_linear_search.pdf}
        \caption{Principle and output of the \textit{linear search}. The simulated dataset (top) corresponds to the one shown and described in \autoref{fig:app_principle_dataset}. First, a set of durations and depths $\set{T_i, D_j}_{i,j}$ is generated. For each pair of indices $(i,j)$, the likelihood $\ln p(\bm{f} \vert T_i ,D_j, \Delta_{i,j})$ is computed using the parameters from \autoref{eq:ls} and the expression of \autoref{eq:linear_search_ll}. This process yields the grid of log-likelihoods $\ln\mathcal{L}$ (bottom plot), as well as the $\set{\Delta_{i,j}, \sigma_{i,j}}_{i, j}$ transit depths and errors inferred linearly using \autoref{eq:ls}.}
        \label{fig:linear_search}
    \end{centering}
\end{figure}

@@

\begin{figure}[H]
    \centering
    \includegraphics[width=0.85\linewidth]{../workflows/principle/figures/principle_periodic_0.pdf}\\
    {(a) $\ln \mathcal{L}$}

    \centering
    \includegraphics[width=0.85\linewidth]{../workflows/principle/figures/principle_periodic_1.pdf}\\
    {(b) $\mathcal{P}(P)$}
\end{figure}

@@

\begin{figure}[H]
    \begin{centering}
        \includegraphics[width=0.85\linewidth]{principle_Q.pdf}
        \caption{For each period $P$, the joint likelihood $\mathcal{P}(P)$ is computed using \autoref{eq:result}, and the value of the maximum likelihood transit SNR retained as $\mathcal{Q}(P)$.}
        \label{fig:periodogram}
    \end{centering}
\end{figure}
